### 2025-01-31 products/manuscript/Manuscript.qmd
Part One Completed
Part One components - description of data set, research question formation, ideas for data cleaning all included in this commit-- also officially obtained my 2019-2024 human tick-bourne illness case data in conjunction with my internship that I am doing! Super excited to proceed.

---
title: "Analysis of Tick-Bourne Disease Cases Reported to DPH Between 2019-2024"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Hope Grismer $^{1,2,*, \land}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2,}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Southeastern Center of Excellence in Vector-Bourne Diseases (SECVBD), Gainesville, FL 32611.
3.  Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: hope.grismer\@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Tick-borne illnesses are diseases that humans can contract from the bite of an infected tick. Ticks are small, blood-sucking arachnids that are often found in wooded, grassy, or suburban areas. When a tick feeds on an infected animal, it can acquire pathogens—such as bacteria, viruses, or parasites—that can then be transmitted to humans. Examples of tick-bourne diseases include, but are not limited to, Lyme disease, anaplasmosis, and babesiosis. These diseases are of increasing public health concern, especially as the geographic range and seasonality of ticks expand due to environmental and ecological changes.

## Description of data and data source

This data set includes all human cases reported to the Georgia Department of Public Health (GA-DPH) between 2019 and 2024. It was graciously provided by DPH professionals in the epidemiology department. All available data for each case over this five-year period across the state of Georgia. Proper de-identification practices-- the removal of first name, last name, and date of birth (DOB)-- has been taken. The dataset includes variables such as disease type, patient demographics (gender, age, district, state, county, and ZIP code), medical details (incident ID, report ID, lab test dates, results, serotype, and titers), as well as disease outcomes (death status, fever, migranes, and administrative status). Reported diagnoses included in this dataset include Toxoplasmosis, Dengue, Rocky Mountain Spotted Fever, Anaplasma Phagocytophilum, Malaria, Lyme Disease, West Nile (WNV) Infection, Chikungunya, LaCrosse (LAC) Infection, Zika, Ehrlichia Chaffeensis, and Typhus (Epidemic/Louse/Flea). Some of this cases, based on adherence to specific case definitions and subsequent lab results, are categorized as suspect, probable, confirmed, or unconfirmed. By examining case reports, public health professionals can better understand the prevalence, demographics, and regional variations of these diseases. This analysis aims to further inform the public health understanding of tick-bourne disease in Georgia.

## Questions/Hypotheses to be addressed

1.  How have vector-borne diseases (VBD) changed over time, and what seasonal patterns or climatic factors influence these trends?

2.  What is the distribution of vector-borne diseases across different health districts, counties, and ZIP codes?

3.  How complete is the data for vector-borne diseases, and how has data completeness evolved over time?

4.  What is the timeline between key dates (onset, lab test, and date of entry), excluding same-day onset and lab results, and how does this timeline vary over time?

5.  What trends exist in malaria travel cases over time, and which countries are most frequently visited by affected individuals?

6.  What are the trends and changes observed in relative incidence of different disease pathologies between 2019-2024?

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

***I. Describing the Uncleaned Data***

The data contains 6255 entries representing discrete cases of tick-bourne diseases. For each case, some or all of the following data variables were collected: disease type, patient demographics (gender, age, district, state, county, and ZIP code), medical details (incident ID, report ID, lab test dates, results, serotype, and titers), and disease outcomes (death status, fever, migranes, and administrative status). For some cases, specific variables were not reported and, therefore, are reported as blank spaces in the data set.

***II. Initial Data Cleaning Stages***

Primary data cleaning anticipated at this initial stage (this section to be rewritten later)

-   Remove all cases that were "NOT CONFIRMED," "UNCONFIRMED," or left blank in the STATUS category

-   Remove entries that contain any blanks to ensure all data analyzed later include complete profiles

-   Sort by Gender, County, and Age

***III. Analysis Process***

-   Still determining this

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References