### 2025-02-21 code/processing-code/processing-code-tickcases.qmd
Part Two Completed
Part Two is completed I believe; Note to self on reference manager and double checking my ReadMe files

---
title: "Data Cleaning for Raw Tick-Bourne Cases Data - 2019 to 2024"
author: "Hope Grismer"
date: "2025-02-21"
output: html_document
---

Data Cleaning for Raw Tick-Bourne Cases Data - 2019 to 2024

```{r}
library(readxl)  # For loading Excel files
library(dplyr)   # For data processing/cleaning
library(tidyr)   # For data processing/cleaning
library(skimr)   # For nice visualization of data
library(here)    # To set paths
library(kableExtra)
```

```{r}
## ---- loaddata --------
# Define the path to the data
data_location <- here::here("data", "raw-data", "Tick-Bourne_Cases_2019_to_2024.xlsx")
```

```{r}
# Load the data
rawdata <- readxl::read_excel(data_location)
```

```{r}
## ---- exploredata --------
# Take a look at the raw data structure
dplyr::glimpse(rawdata)
head(rawdata)
```

```{r}
# Summary statistics
summary(rawdata)

# Preview first few rows
head(rawdata)

# Nice visualization of data
skimr::skim(rawdata)
```
The first cleaning step removes cases where the STATUS field is either missing or not labeled as "CONFIRMED." This ensures that only officially confirmed cases are included in further analysis. The cleaned dataset is then saved for future reference.

```{r}
## ---- cleandata --------
# Remove cases where STATUS is NA or not confirmed
cleaned_data <- rawdata %>%
  dplyr::filter(!is.na(STATUS) & STATUS == "CONFIRMED")

# Take another look at cleaned data
dplyr::glimpse(cleaned_data)

## ---- savedata --------
# Define save location for cleaned data
save_data_location <- here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned.rds")

# Ensure the "processed-data" folder exists
if (!dir.exists(here::here("data", "processed-data"))) {
  dir.create(here::here("data", "processed-data"), recursive = TRUE)
}
# Save the cleaned data
saveRDS(cleaned_data, file = save_data_location)

# Confirm success
message("Cleaned data saved successfully at: ", save_data_location)
```

```{r}
## ---- Table --------
cleaned_data %>%
  dplyr::select(DISEASE, AGE, GENDER, COUNTY, STATE, LABDATE1, RESULT1, STATUS) %>%  # Select key columns for display
  knitr::kable(caption = "Cleaned Tick-Bourne Disease Cases (2019-2024)", align = "c") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  ) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight the first column
  kableExtra::column_spec(2:8, width = "12em") %>%  # Adjust column widths for readability
  kableExtra::row_spec(0, bold = TRUE, background = "#34495E", color = "white")  # Style header row
```
To refine the dataset further, only cases with a confirmed "positive" test result are retained. Any cases with missing, blank, or ambiguous results (such as "inconclusive") are excluded. This step eliminated 15 cases that did not meet the case definition used in Georgia, leaving 95 valid data points. However, upon further review, I noticed that additional positive cases that were not explicitly labeled "positive" were inadvertently removed, prompting a refinement of the filtering criteria.

```{r}
## ---- clean_data_step2 --------
# Further exclude cases where RESULT1 is blank or not "positive" in any capitalization
cleaned_data2 <- cleaned_data %>%
  filter(!is.na(RESULT1) & RESULT1 != "" & tolower(RESULT1) == "positive")

# Save final cleaned dataset
final_save_data_location <- here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned_Step2.rds")
saveRDS(cleaned_data2, file = final_save_data_location)

## ---- display_table_step2 --------
# Display professional table for second cleaning step
kableExtra::kable(cleaned_data2, format = "html", caption = "Final Cleaned Data After Removing Blank RESULT1 Cases") %>%
  kableExtra::kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight first column
  kableExtra::column_spec(2:ncol(cleaned_data2), width = "10em") %>%  # Adjust column widths
  kableExtra::row_spec(0, bold = TRUE, background = "#34495E", color = "white")  # Header row styling
```
This step provided an easier summary table of some of the most relevant information on these positive cases.

```{r}
## ---- display only a few of the columns --------

# Select the specified columns and sort by DISEASE
selected_columns_data <- cleaned_data2 %>%
  select(DISEASE, STATUS, COUNTY, GENDER, AGE, LABTEST1, RESULT1) %>%
  dplyr::arrange(DISEASE)  # Ensure sorting by disease name

# Display the table with consistent styling
kableExtra::kable(selected_columns_data, format = "html", caption = "Selected Columns from Final Cleaned Data") %>%
  kableExtra::kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight the first column
  kableExtra::column_spec(1:ncol(selected_columns_data), width = "12em") %>%  # Adjust column widths
  kableExtra::row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")  # Header row styling
```
A second round of filtering ensures that all positive cases are correctly identified while excluding known non-positive values such as "Negative," "Equivocal," "Not detected," and other irrelevant entries. The updated dataset is saved again for consistency.
```{r}
## ---- clean_data_step3--------
# Define values to exclude from RESULT1
excluded_values <- c("", NA, "Negative", "Equivocal", "negative", "Choose One", "Not detected", "Indeterminant/Equivocal")

# Filter out only the excluded values while keeping all other results
cleaned_data3 <- cleaned_data %>%
  filter(!(RESULT1 %in% excluded_values) & !is.na(RESULT1))

# Save final cleaned dataset
cleaned_data3_save_location <- here::here("data", "processed-data", "Tick-Bourne_Cases_Final_Cleaned.rds") 
saveRDS(cleaned_data3, file = cleaned_data3_save_location)

# Display count of remaining observations
nrow(cleaned_data3)

```
To present the cleaned data effectively, selected key columns—including DISEASE, STATUS, COUNTY, GENDER, AGE, LABTEST1, and RESULT1—are displayed in a professionally styled table using the kableExtra package. This enhances readability and makes it easier to analyze trends.
```{r}
## ---- display only a few of the columns --------

# Select the specified columns
selected_columns_data <- cleaned_data3 %>%
  select(DISEASE, STATUS, COUNTY, GENDER, AGE, LABTEST1, RESULT1) %>%
  dplyr::arrange(DISEASE)  # Use the correct column name

# Display the table with consistent styling
kableExtra::kable(selected_columns_data, format = "html", caption = "Selected Columns from Final Cleaned Data") %>%
  kableExtra::kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight the first column
  kableExtra::column_spec(1:ncol(selected_columns_data), width = "12em") %>%  # Adjust column widths
  kableExtra::row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")  # Header row styling

```
A histogram is created to visualize the distribution of different tick-borne diseases in Georgia between 2019 and 2024. The ggplot2 package is used to generate a clean and informative bar chart, illustrating the most prevalent diseases. 
```{r}
## ---- disease_histogram --------
library(ggplot2)

# Create histogram of disease counts
disease_plot <- ggplot(cleaned_data3, aes(x = DISEASE)) +
  geom_bar(fill = "#2C3E50", color = "white") +  # Dark blue bars with white edges
  theme_minimal() +
  labs(
    title = "Distribution of Tick-Bourne Disease\nTypes for Confirmed Cases in Georgia (2019-2024)",
    x = "Disease Type",
    y = "Count"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),  # Rotate & reduce x-axis label size
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.margin = margin(10, 10, 20, 10)  # Adjust margins to prevent cutoff
  )

# Display the plot
print(disease_plot)

# Define save locations
save_plot_location <- here::here("results", "figures", "disease_histogram.png")

# Ensure directory exists
dir.create(here::here("results", "figures"), recursive = TRUE, showWarnings = FALSE)

# Save the plot
ggsave(filename = save_plot_location, plot = disease_plot, width = 12, height = 6, dpi = 300)
```
To explore the severity of Rocky Mountain Spotted Fever cases, a box plot of TITER1 values (log-transformed for better visualization) is created. This helps in understanding the spread and variability of test results among confirmed cases. It also is considered relatively analagous to the patterns seen in Alpha-Gal Syndrome, which is a topic of interest for me, but is not explicitly included in this data set as it is currently not a reportable disease in Georgia.
```{r}
# Save the plot
ggsave(filename = save_plot_location, plot = disease_plot, width = 12, height = 6, dpi = 300)

# ---- rmsf_boxplot --------
# Filter for Rocky Mountain Spotted Fever cases
rmsf_data <- rawdata %>%
  filter(DISEASE == "ROCKY MOUNTAIN SPOTTED FEVER") %>%
  mutate(TITER1 = as.numeric(TITER1)) %>%  # Ensure TITER1 is numeric
  filter(!is.na(TITER1) & TITER1 > 0)

# Create box and whisker plot for log-transformed TITER1 values
rmsf_titer_plot <- ggplot(rmsf_data, aes(y = TITER1)) +
  geom_boxplot(fill = "#1B9E77", color = "black") +  # Green box plot
  scale_y_log10() +  # Apply log10 scale
  theme_minimal() +
  labs(
    title = "Log-Transformed Distribution of TITER1 Values\nfor Rocky Mountain Spotted Fever Cases",
    y = "TITER1 Value (Log Scale)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank(),   # Remove x-axis labels
    axis.ticks.x = element_blank()   # Remove x-axis ticks
  )

# Display the plot
print(rmsf_titer_plot)

# Define save location
save_rmsf_plot_location <- here::here("results", "figures", "rmsf_titer_boxplot_log.png")

# Save the plot
ggsave(filename = save_rmsf_plot_location, plot = rmsf_titer_plot, width = 8, height = 6, dpi = 300)

```
The dataset is filtered to determine the five most common tick-borne diseases. A bar chart is created to highlight the case count for each of these diseases, providing an overview of the most significant public health concerns.
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(here)

# Load the cleaned data
cleaned_data <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned.rds"))

# Get top 5 diseases by total count across all counties
top_diseases <- cleaned_data %>%
  count(DISEASE, sort = TRUE) %>%  # Count cases per disease
  top_n(5, n) %>%  # Select top 5 diseases
  pull(DISEASE)  # Extract disease names as a vector

# Filter data for only the top 5 diseases
filtered_data <- cleaned_data %>%
  filter(DISEASE %in% top_diseases)

# Summarize total cases for each of the top 5 diseases
disease_summary <- filtered_data %>%
  group_by(DISEASE) %>%
  summarise(Count = n(), .groups = "drop")  # Count cases per Disease

# Create bar chart
top5_disease_plot <- ggplot(disease_summary, aes(x = reorder(DISEASE, -Count), y = Count, fill = DISEASE)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8) +  # Bar chart for count
  geom_text(aes(label = Count), vjust = -0.5, size = 5) +  # Count labels above bars
  theme_minimal() +
  labs(
    title = "Top 5 Tick-Borne Diseases",
    x = "Disease",
    y = "Case Count",
    fill = "Disease"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10)  # Angled labels for clarity
  ) +
  scale_fill_brewer(palette = "Set1")  # Nice color scheme for different diseases

# Display the plot
print(top5_disease_plot)

# Define save location
save_top5_disease_plot <- here::here("results", "figures", "top5_diseases_plot.png")

# Save the plot
ggsave(filename = save_top5_disease_plot, plot = top5_disease_plot, width = 8, height = 6, dpi = 300)

# Confirm success
message("Plot saved successfully at: ", save_top5_disease_plot)

```
Lyme Disease is likewise similar to Alpha-gal. So, a specific analysis is conducted for Lyme Disease cases, focusing on test results was conducted. Below ia a count histogram generated to compare positive, negative, and other test outcomes, offering insights into testing patterns and diagnosis rates.

```{r}
# ---- lyme_count_histogram --------
# Filter for Lyme Disease cases and exclude "Choose One" category
lyme_data2 <- rawdata %>%
  filter(DISEASE == "LYME DISEASE", RESULT1 != "Choose One") %>%
  mutate(
    RESULT1 = case_when(
      RESULT1 %in% c("Unknown", "Other") ~ "Unknown/Other",  # Group "Unknown" and "Other"
      RESULT1 == "Results Pending" ~ "Results Pending",      # Keep "Results Pending" separate
      TRUE ~ RESULT1  # Keep all other values unchanged
    ),
    RESULT1 = factor(RESULT1)  # Ensure it's a factor
  )

# Create count histogram for positive vs. negative cases
lyme_count_histogram2 <- ggplot(lyme_data2, aes(x = RESULT1, fill = RESULT1)) +
  geom_bar(color = "black", alpha = 0.8) +  # Bar chart for count comparison
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5, size = 5) +  # Count labels above bars
  scale_fill_manual(values = c(
    "Positive" = "#E41A1C",
    "Negative" = "#377EB8",
    "Unknown/Other" = "#4DAF4A",   # Green for the grouped category
    "Results Pending" = "#FF7F00"  # Orange for "Results Pending"
  )) +
  theme_minimal() +
  labs(
    title = "Count of Lyme Disease Test Results",
    x = "Test Result",
    y = "Count",
    fill = "Test Result"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(face = "bold", size = 8, angle = 45, hjust = 1)  # Smaller labels
  )

# Display the plot
print(lyme_count_histogram2)

# Define save location
save_lyme_count_plot <- here::here("results", "figures", "lyme_result_count_histogram2.png")

# Save the plot
ggsave(filename = save_lyme_count_plot, plot = lyme_count_histogram2, width = 8, height = 6, dpi = 300)

```

The final visualization maps tick-borne disease cases across different counties in Georgia. A grouped bar chart displays case counts per county, categorized by disease type. This geographic breakdown helps in identifying regional hotspots for tick-borne illnesses, which can inform public health interventions. This visualization is busy so it was further altered to show the top five tick bourne diseases and how they break down across the top 12 (by overall count of confirmed cases) counties in Georgia.

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(here)

# Load the cleaned data
cleaned_data <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned.rds"))

# Create count summary by Disease and County
disease_county_summary <- cleaned_data %>%
  group_by(DISEASE, COUNTY) %>%
  summarise(Count = n(), .groups = "drop")  # Count cases per Disease & County

# Create bar chart
disease_county_plot <- ggplot(disease_county_summary, aes(x = COUNTY, y = Count, fill = DISEASE)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8) +  # Bar chart for count
  geom_text(aes(label = Count), vjust = -0.5, size = 3) +  # Count labels above bars
  theme_minimal() +
  labs(
    title = "Tick-Borne Disease Cases by County",
    x = "County",
    y = "Case Count",
    fill = "Disease"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8)  # Angled labels for clarity
  ) +
  scale_fill_brewer(palette = "Set1")  # Nice color scheme for different diseases

# Display the plot
print(disease_county_plot)

# Define save location
save_disease_county_plot <- here::here("results", "figures", "disease_county_plot.png")

# Save the plot
ggsave(filename = save_disease_county_plot, plot = disease_county_plot, width = 10, height = 6, dpi = 300)

# Confirm success
message("Plot saved successfully at: ", save_disease_county_plot)

```
To look at trends in disease burden as well as "hot-spot" counties, I generated the stacked bar graph below.
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(here)
library(RColorBrewer)

# Load the cleaned data
cleaned_data <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned.rds"))

# Get top 5 diseases by total count across all counties
top_diseases <- cleaned_data %>%
  count(DISEASE, sort = TRUE) %>%  
  top_n(5, n) %>%  
  pull(DISEASE)

# Filter data for only the top 5 diseases
filtered_data <- cleaned_data %>%
  filter(DISEASE %in% top_diseases)

# Identify the top 12 counties by total case count
top_counties <- filtered_data %>%
  count(COUNTY, sort = TRUE) %>%
  top_n(12, n) %>%
  pull(COUNTY)

# Ensure we have exactly 12 counties (or fewer if dataset is small)
top_counties <- unique(top_counties)[1:min(12, length(top_counties))]

# Keep only those top counties in the dataset
filtered_data <- filtered_data %>%
  filter(COUNTY %in% top_counties)

# Summarize case counts for each disease and county
disease_county_summary <- filtered_data %>%
  group_by(DISEASE, COUNTY) %>%
  summarise(Count = n(), .groups = "drop")

# Get the exact number of unique counties in filtered data
num_counties <- length(unique(disease_county_summary$COUNTY))

# Assign colors dynamically based on available counties
county_colors <- setNames(brewer.pal(num_counties, "Paired"), unique(disease_county_summary$COUNTY))

# Create stacked bar chart
top5_stacked_plot <- ggplot(disease_county_summary, aes(x = DISEASE, y = Count, fill = COUNTY)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8) +  
  theme_minimal() +
  labs(
    title = "Top 5 Tick-Borne Diseases by County",
    x = "Disease",
    y = "Case Count",
    fill = "County"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10)
  ) +
  scale_fill_manual(values = county_colors)  # Assign fixed colors to the top counties

# Display the plot
print(top5_stacked_plot)

# Define save location
save_top5_stacked_plot <- here::here("results", "figures", "top5_diseases_stacked_plot.png")

# Save the plot
ggsave(filename = save_top5_stacked_plot, plot = top5_stacked_plot, width = 8, height = 6, dpi = 300)

# Confirm success
message("Plot saved successfully at: ", save_top5_stacked_plot)

```
Age distribution might be of interest in the analysis, so this was checked as part of this initial data cleaning. This output shows 0 rows of the cleaned data to be missing ages. As an exploratory analysis, I looked at disease count by age group from my cleaned data set. 
```{r}
missing_age_rows <- cleaned_data3 %>%
  filter(is.na(AGE))

print(missing_age_rows)

library(ggplot2)
library(dplyr)

# Create age bins of 20 years
selected_columns_data <- cleaned_data2 %>%
  select(DISEASE, AGE) %>%
  mutate(Age_Bin = cut(AGE, 
                       breaks = seq(0, 100, by = 20),  # Adjust max age as needed
                       labels = c("0-19", "20-39", "40-59", "60-79", "80+"),
                       right = FALSE)) %>%
  drop_na(Age_Bin)  # Remove any NA values due to missing AGE data

# Create stacked bar plot
age_group_plot <- ggplot(selected_columns_data, aes(x = Age_Bin, fill = DISEASE)) +
  geom_bar(position = "stack") +
  labs(title = "Disease Count by Age Group",
       x = "Age Group",
       y = "Count",
       fill = "Disease Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

# Save the plot to results/figures using here::here()
ggsave(
  filename = here::here("results", "figures", "disease_count_by_age.png"), 
  plot = age_group_plot, 
  width = 8, 
  height = 6, 
  dpi = 300
)

age_group_plot

```
Another thing that might be used in my analysis to answer the question of how data completeness evolved over time is cleaning the data to extract data onset information.

By distinguishing between Blank and Text in the STATUS column over time, this plot helps pinpoint periods with high missingness. If there are spikes in Blank values at certain times, it could indicate data entry issues, system errors, or gaps in reporting that need to be addressed before deeper analysis.
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)

# Convert DOO to Date and group by Month
statuscompleteness_summary <- rawdata %>%
  mutate(DOO = as.Date(DOO),  # Ensure DOO is in Date format
         Month = floor_date(DOO, "month"),  # Aggregate by month
         Status_Category = ifelse(STATUS == "" | is.na(STATUS), "Blank", "Text")) %>%
  group_by(Month, Status_Category) %>%
  summarise(Count = n(), .groups = "drop")

# Save the cleaned dataset to the processed_data folder
write_csv(statuscompleteness_summary, here::here("data", "processed-data", "cleaned_statuscompleteness_summary.csv"))

# Create improved line plot
status_completeness_plot <- ggplot(statuscompleteness_summary, aes(x = Month, y = Count, color = Status_Category, group = Status_Category)) +
  geom_line(size = 1) +  # Adjust line thickness
  geom_point(size = 1, alpha = 0.6) +  # Reduce point size & add transparency
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed") +  # Add smoothed trendline
  labs(title = "Comparison of Blank vs. Text in STATUS Column by Month",
       x = "Date of Onset (Monthly Aggregation)",
       y = "Count",
       color = "Status Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Save the plot to results/figures using here::here()
ggsave(
  filename = here::here("results", "figures", "status_completeness_plot.png"), 
  plot = status_completeness_plot, 
  width = 8, 
  height = 6, 
  dpi = 300
)
status_completeness_plot

```

```{r}
library(dplyr)
library(lubridate)
library(readr)
library(here)
library(kableExtra)
library(fs)

# Extract year and month from the selected date columns
processed_date_data <- rawdata %>%
  mutate(
    UPDATE_YEAR = year(ymd(UPDATEDATE)),
    UPDATE_MONTH = month(ymd(UPDATEDATE), label = TRUE),  # Extract month name
    DOO_YEAR = year(ymd(DOO)),
    DOO_MONTH = month(ymd(DOO), label = TRUE),
    LAB1_YEAR = year(ymd(LABDATE1)),
    LAB1_MONTH = month(ymd(LABDATE1), label = TRUE),
    LAB2_YEAR = year(ymd(LABDATE2)),
    LAB2_MONTH = month(ymd(LABDATE2), label = TRUE)
  ) %>%
  filter(
    !is.na(UPDATE_YEAR) & !is.na(UPDATE_MONTH) & 
    !is.na(DOO_YEAR) & !is.na(DOO_MONTH) & 
    !is.na(LAB1_YEAR) & !is.na(LAB1_MONTH) & 
    !is.na(LAB2_YEAR) & !is.na(LAB2_MONTH)  # Exclude rows with missing date parts
  )

# Select key columns for review
summary_table <- processed_date_data %>%
  select(DISEASE, UPDATEDATE, DOO, LABDATE1, LABDATE2)

# Save the cleaned dataset
fs::dir_create(here::here("data", "processed-data"))  # Ensure directory exists
write_csv(processed_date_data, here::here("data", "processed-data", "processed_dates.csv"))

# Summary Table
kableExtra::kable(summary_table, format = "html", caption = "Extracted Date Information for Disease Cases") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight the first column
  kableExtra::column_spec(2:ncol(summary_table), width = "12em") %>%  # Adjust column widths
  kableExtra::row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")  # Header row styling

```
Another question that I want to analyze later and therefore needed to clean for was What is the timeline between key dates (onset, lab test, and date of entry), excluding same-day onset and lab results, and how does this timeline vary over time? One thing I have noticed in my internship work on tick-bourne illness is the huge issue with loss-to-follow-up.

The code below calculate the time differences between key dates. The script then calculates the number of days between the onset date (DOO) and the first lab date (LABDATE1), between the first and second lab dates (LABDATE2), and between the second lab date and the update date (UPDATEDATE). These values are summed to determine the total duration of each case.

This processed dataset is then saved as a CSV file for further analysis. The script also generates a formatted HTML table to display the extracted time information in a structured manner.

Once I took a look at this table, however, there were big issues. Namely, that negative time (in days) was recorded between the initial lab result and the follow up test. I wrote further code to remove these values because it is not able to be determined what the proper date was due to user error. I added a filter accordingly.

```{r}
library(dplyr)
library(lubridate)
library(readr)
library(here)
library(kableExtra)
library(fs)
library(ggplot2)

# Calculate time differences between the dates
processed_time_data <- processed_date_data %>%
  mutate(
    DOO_to_LAB1 = as.numeric(difftime(LABDATE1, DOO, units = "days")),
    LAB1_to_LAB2 = as.numeric(difftime(LABDATE2, LABDATE1, units = "days")),
    LAB2_to_UPDATE = as.numeric(difftime(UPDATEDATE, LABDATE2, units = "days")),
    TOTAL_DAYS = DOO_to_LAB1 + LAB1_to_LAB2 + LAB2_to_UPDATE
  )

# Remove rows with any negative values
processed_time_data2 <- processed_time_data %>%
  filter(if_all(c(DOO_to_LAB1, LAB1_to_LAB2, LAB2_to_UPDATE, TOTAL_DAYS), ~ . >= 0))

# Select key columns for review
time_summary_table <- processed_time_data2 %>%
  select(DISEASE, DOO, LABDATE1, LABDATE2, UPDATEDATE, DOO_to_LAB1, LAB1_to_LAB2, LAB2_to_UPDATE, TOTAL_DAYS)

# Save the dataset with time differences
fs::dir_create(here::here("data", "processed-data"))  # Ensure directory exists
write_csv(processed_time_data2, here::here("data", "processed-data", "time_between_processeddata.csv"))

# Summary Table
kableExtra::kable(time_summary_table, format = "html", caption = "Time Differences Between Key Date Points") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::column_spec(1, bold = TRUE, color = "white", background = "#2C3E50") %>%  # Highlight the first column
  kableExtra::column_spec(2:ncol(time_summary_table), width = "12em") %>%  # Adjust column widths
  kableExtra::row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")  # Header row styling

```