### 2025-03-17 code/analysis-code/statistical-analysis.qmd
Part 3 Done
Made some better manuscript edits

---
title: "Data Analysis for Raw Tick-Bourne Cases Data - 2019 to 2024"
author: "Hope Grismer"
date: "2025-03-10"
output: html_document
---

######################################
# analysis script
#
# This script loads the processed, cleaned data, does a simple analysis
# and saves the results to the results folder.

First, load all the libraries needed for tidymodels, multivariable GLM, and data visualization.

```{r}
library(tidymodels)  # For tidymodels framework
library(ggplot2)     # For data visualization
library(dplyr)       # For data manipulation
library(lubridate)   # For date manipulation
library(readr)       # For reading and writing CSV files
library(fs)          # For file system operations
library(broom)       # For tidying up GLM results
library(here)        # for data loading/saving
```

Then, load the cleaned datasets.

```{r}
# Load the cleaned dataset
cleaned_data2 <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned_Step2.rds"))
processed_time_data2 <- read_csv(here::here("data", "processed-data", "time_between_processeddata.csv"))

```

While the processing code creation involved some exploratory data analysis (EDA) already to help me determine what cleaning needed to be done, additional EDA helped find patterns in data to analyze further. See below.

```{r}

# Visualize time differences (e.g., DOO to LAB1)
ggplot(processed_time_data2, aes(x = DOO_to_LAB1)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Time from DOO to LAB1 (in days)", 
       x = "Days from DOO to LAB1", y = "Frequency")

```

######################################
# Data fitting/statistical analysis
#####################################

```{r}
# Path to the cleaned data (after saving it)
cleaned_data3 <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Final_Cleaned.rds"))
```

```{r}
# 1. Disease trends over time:
# Extract YEAR from DOO (assuming DOO is a Date column)
cleaned_data3 <- cleaned_data3 %>%
  mutate(YEAR = as.integer(format(as.Date(DOO, format = "%Y-%m-%d"), "%Y")))

# Count cases grouped by DISEASE type and YEAR
disease_counts <- cleaned_data3 %>%
  group_by(DISEASE, YEAR) %>%
  summarise(CASE_COUNT = n(), .groups = "drop")

# Fit a linear model for yearly trend in cases (by DISEASE)
lmfit1 <- lm(CASE_COUNT ~ YEAR + DISEASE, data = disease_counts)
lmtable1 <- broom::tidy(lmfit1)
print(lmtable1)

# Save fit results
table_file1 <- here::here("results", "tables", "trend_over_time_by_disease.rds")
saveRDS(lmtable1, file = table_file1)

# Plot yearly trend by disease
plot1 <- ggplot(disease_counts, aes(x = YEAR, y = CASE_COUNT, color = DISEASE)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm") +
  labs(
    title = "Vector-Borne Diseases Over Time by Disease",
    x = "Year",
    y = "Number of Cases"
  ) +
  theme_minimal()

plot1

# Save plot
plot_file1 <- here::here("results", "figures", "trend_over_time_by_disease.png")
ggsave(plot_file1, plot = plot1, width = 8, height = 6, dpi = 300)
```

```{r}
# 2. Distribution Across Health Districts:
# Bar plot of cases by health district
plot2 <- ggplot(cleaned_data3, aes(x = COUNTY, fill = DISEASE)) +
  geom_bar() +
  labs(
    title = "Cases by Health District",
    x = "Health District",
    y = "Number of Cases"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot2

# Save plot
plot_file2 <- here::here("results", "figures", "cases_by_health_district.png")
ggsave(plot_file2, plot = plot2, width = 8, height = 6, dpi = 300)
```

```{r}
# 3. Completeness 
library(ggplot2)
library(dplyr)
library(lubridate)
library(broom)  # for tidying model results
library(readr)  # for reading CSV files

# Load the cleaned data (adjust path as needed)
statuscompleteness_summary_location <- here::here("data", "processed-data", "cleaned_statuscompleteness_summary.csv")
statuscompleteness_summary <- read_csv(statuscompleteness_summary_location)

# 1. Data Completeness Analysis Over Time:
# Summarize completeness over time
completeness_summary <- statuscompleteness_summary %>%
  mutate(YEAR = year(Month)) %>%
  group_by(YEAR) %>%
  summarise(COMPLETENESS = mean(Count) / sum(Count) * 100, .groups = "drop")  # Calculate completeness as percentage

# Linear Model: Analyze the relationship between YEAR and COMPLETENESS
lm_completeness <- lm(COMPLETENESS ~ YEAR, data = completeness_summary)

# Summarize the linear model results
lm_completeness_summary <- broom::tidy(lm_completeness)
print(lm_completeness_summary)

# 2. Create the completeness plot with linear trendline
plot3 <- ggplot(completeness_summary, aes(x = YEAR, y = COMPLETENESS)) +
  geom_line(color = "darkgreen") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +  # Add linear model trendline
  labs(
    title = "Data Completeness Over Time",
    x = "Year",
    y = "Completeness (%)"
  ) +
  theme_minimal()

# Print the plot
plot3

# Save the plot to the results/figures folder
plot_file3 <- here::here("results", "figures", "completeness_over_time.png")
ggsave(plot_file3, plot = plot3, width = 8, height = 6, dpi = 300)

```

```{r}
# 4. Timeline Between Key Dates:
# Fit a model for days between onset and lab over time
filtered_time_data <- cleaned_data3 %>%
  filter(!is.na(LABDATE1) & !is.na(DOO) & (LABDATE1 > DOO))

# Calculate the days between onset and lab result (assuming DATEONSET and LABDATE1 are in Date format)
filtered_time_data <- filtered_time_data %>%
  mutate(DAYS_TO_LAB = as.integer(LABDATE1 - DOO))

lmfit2 <- lm(DAYS_TO_LAB ~ YEAR, data = filtered_time_data)
lmtable2 <- broom::tidy(lmfit2)
print(lmtable2)

# Save fit results
table_file2 <- here::here("results", "tables", "days_to_lab_over_time.rds")
saveRDS(lmtable2, file = table_file2)

# Plot days between onset and lab over time
plot4 <- ggplot(filtered_time_data, aes(x = YEAR, y = DAYS_TO_LAB)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Days Between Onset and Lab Processing Over Time",
    x = "Year",
    y = "Days Between Onset and Lab"
  ) +
  theme_minimal()

plot4

# Save plot
plot_file4 <- here::here("results", "figures", "days_to_lab_over_time.png")
ggsave(plot_file4, plot = plot4, width = 8, height = 6, dpi = 300)

```

```{r}
# 5. Relative Incidence of Disease Pathologies (2019–2024):
# Filter data for years 2019–2024
incidence_data <- cleaned_data3 %>%
  filter(YEAR >= 2019 & YEAR <= 2024)

# Group and calculate incidence by disease
incidence_summary <- incidence_data %>%
  group_by(YEAR, DISEASE) %>%
  summarise(INCIDENCE = n()) %>%
  ungroup()

# Plot relative incidence over time
plot5 <- ggplot(incidence_summary, aes(x = YEAR, y = INCIDENCE, color = DISEASE)) +
  geom_line(size = 1) +
  labs(
    title = "Relative Incidence of Disease Pathologies (2019–2024)",
    x = "Year",
    y = "Incidence"
  ) +
  theme_minimal()

plot5

# Save plot
plot_file5 <- here::here("results", "figures", "relative_incidence_2019_2024.png")
ggsave(plot_file5, plot = plot5, width = 8, height = 6, dpi = 300)
```