### 2025-03-17 products/manuscript/Manuscript.qmd

---
title: "Analysis of Tick-Bourne Disease Cases Reported to DPH Between 2019-2024"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

The structure below is one possible setup for a manuscript, or a general
data analysis project (including the course project). Adjust as needed.
You don't need to have exactly these sections, but the content covering
those sections should be addressed.

This uses MS Word as output format. [See
here](https://quarto.org/docs/output-formats/ms-word.html) for more
information. You can switch to other formats, like html or pdf. See [the
Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Hope Grismer $^{1,2,*, \land}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2,}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Southeastern Center of Excellence in Vector-Bourne Diseases
    (SECVBD), Gainesville, FL 32611.
3.  Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: hope.grismer\@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Tick-borne illnesses are diseases that humans can contract from the bite
of an infected tick. Ticks are small, blood-sucking arachnids that are
often found in wooded, grassy, or suburban areas. When a tick feeds on
an infected animal, it can acquire pathogens—such as bacteria, viruses,
or parasites—that can then be transmitted to humans. Examples of
tick-bourne diseases include, but are not limited to, Lyme disease,
Anaplasmosis, and babesiosis. These diseases are of increasing public
health concern, especially as the geographic range and seasonality of
ticks expand due to environmental and ecological changes.

## Description of data and data source

This data set includes all human cases reported to the Georgia
Department of Public Health (GA-DPH) between 2019 and 2024. It was
graciously provided by DPH professionals in the epidemiology department.
All available data for each case over this five-year period across the
state of Georgia. Proper de-identification practices-- the removal of
first name, last name, and date of birth (DOB)-- has been taken. The
dataset includes variables such as disease type, patient demographics
(gender, age, district, state, county, and ZIP code), medical details
(incident ID, report ID, lab test dates, results, serotype, and titers),
as well as disease outcomes (death status, fever, migranes, and
administrative status). Reported diagnoses included in this dataset
include Toxoplasmosis, Dengue, Rocky Mountain Spotted Fever, Anaplasma
Phagocytophilum, Malaria, Lyme Disease, West Nile (WNV) Infection,
Chikungunya, LaCrosse (LAC) Infection, Zika, Ehrlichia Chaffeensis, and
Typhus (Epidemic/Louse/Flea). Some of this cases, based on adherence to
specific case definitions and subsequent lab results, are categorized as
suspect, probable, confirmed, or unconfirmed. By examining case reports,
public health professionals can better understand the prevalence,
demographics, and regional variations of these diseases. This analysis
aims to further inform the public health understanding of tick-bourne
disease in Georgia.

## Questions/Hypotheses to be addressed

1.  How have vector-borne diseases (VBD) changed over time, and what
    seasonal patterns or climatic factors influence these trends?

2.  What is the distribution of vector-borne diseases across different
    health districts, counties, and ZIP codes?

3.  How complete is the data for vector-borne diseases, and how has data
    completeness evolved over time?

4.  What is the timeline between key dates (onset, lab test, and date of
    entry), excluding same-day onset and lab results, and how does this
    timeline vary over time?

5.  What trends exist in malaria travel cases over time, and which
    countries are most frequently visited by affected individuals?

6.  What are the trends and changes observed in relative incidence of
    different disease pathologies between 2019-2024?

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above and have the right bibtex key. Then
you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

**I. Uncleaned Data**

The data contains 6255 entries representing discrete cases of
tick-bourne diseases. For each case, some or all of the following data
variables were collected: disease type, patient demographics (gender,
age, district, state, county, and ZIP code), medical details (incident
ID, report ID, lab test dates, results, serotype, and titers), and
disease outcomes (death status, fever, migranes, and administrative
status). For some cases, specific variables were not reported and,
therefore, are reported as blank spaces in the data set.

***II. Initial Data Cleaning Stages***

**Looked for Missing Data & General Trends**

-   Excluded cases with a missing or non-“CONFIRMED” STATUS to retain
    only confirmed cases.
-   Refined the dataset by retaining only confirmed “positive” test
    results, excluding missing or ambiguous results, leaving 95 valid
    cases.
-   Additional filtering ensured correct identification of positive
    cases while excluding non-positive values.

**Identified Disease Burden Trends & Hotspot Counties**

-   Generated a histogram to visualize the distribution of tick-borne
    diseases in Georgia between 2019 and 2024.
-   Created a bar chart highlighting the five most common diseases.
-   Specifically analyzed Lyme Disease cases, focusing on test results.

**Checked Age Distribution**

-   Verified that no cases had missing age values in the cleaned
    dataset.
-   Conducted an exploratory analysis of disease counts by age group.

**Evaluated Data Completeness Over Time**

-   Extracted onset date information to track missing data trends.
-   Created a plot distinguishing between blank and non-blank STATUS
    entries over time to identify periods of high missingness,
    indicating possible data entry errors or reporting gaps.

**Analyzed Timeline Between Key Case Events**

-   Calculated time differences for:
    -   Onset date (DOO) to first lab test date (LABDATE1)
    -   First lab test (LABDATE1) to second lab test (LABDATE2)
    -   Second lab test (LABDATE2) to case update date (UPDATEDATE)
-   Summed these values to determine the total case duration.
-   Saved the processed dataset as a CSV file for further analysis.
-   Created a formatted HTML table to present the extracted time
    intervals.

**Addressed Data Quality Issues**

-   Identified negative time values between lab test dates, indicating
    data entry errors.
-   Implemented a filtering step to remove records with invalid date
    sequences.

***III. Analysis Process***

-   Still determining this

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not
created with code (if you can do it with code, do it). @fig-schematic is
an example of some - completely random/unrelated - schematic that was
generated with Biorender. We store those figures in the `assets` folder.

## Data aquisition

The data used in this analysis was obtained through my internship with
the Georgia Department of Health, in collaboration with the Centers for
Disease Control and Prevention (CDC). Specifically, it is derived from
the Nationally Notifiable Diseases Surveillance System (NNDSS) and
focuses on tick-borne illnesses. This dataset provides key insights into
the trends, distribution, and characteristics of reported cases,
supporting public health efforts in disease monitoring and prevention.

## Data import and cleaning

The data cleaning and preprocessing workflow begins with loading
essential R packages for data handling (dplyr, tidyr, readxl), summary
visualization (skimr), and file path management (here). The raw dataset,
stored in an Excel file, is imported into R for structured processing.
To assess data quality and structure, we utilize functions such as
glimpse(), head(), and summary(), which provide an overview of column
names, data types, and potential inconsistencies like missing or
unexpected values. Additionally, skimr::skim() offers a detailed
summary, highlighting numerical distributions and missing data patterns.

The initial cleaning step removes records where the STATUS field is
either missing or not classified as "CONFIRMED," ensuring that only
officially verified cases are included. The dataset is then refined
further by retaining only cases with a confirmed "positive" test result,
excluding records with missing, blank, or ambiguous values (e.g.,
"inconclusive"). This process initially removed 15 cases that did not
meet the case definition in Georgia, leaving 95 valid records. Upon
further review, filtering criteria were adjusted to ensure all relevant
positive cases were retained while systematically excluding non-positive
values such as "Negative," "Equivocal," "Not detected," and other
irrelevant categories. This refinement resulted in a more accurate
dataset and a clearer summary table of key case information. The cleaned
dataset is saved at each stage to maintain version control and
reproducibility.

## Key Variables in Data Wrangling - Preparing for Analysis

To effectively present the cleaned dataset, key variables—including
DISEASE, STATUS, COUNTY, GENDER, AGE, LABTEST1, and RESULT1—are
displayed in a professionally formatted table using the kableExtra
package. This enhances readability and facilitates trend analysis.
Several visualizations are generated using ggplot2 to explore patterns
in tick-borne diseases in Georgia between 2019 and 2024: Disease
Distribution: A histogram illustrates the prevalence of different
tick-borne diseases, highlighting the most commonly reported cases.

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your
data. Show the most important descriptive results here. Additional ones
should go in the supplement. Even more can be in the R and Quarto files
that are part of your project.*

Note the loading of the data providing a **relative** path using the
`../../` notation. (Two dots means a folder up). You never want to
specify an **absolute** path like `C:\ahandel\myproject\results\`
because if you share this with someone, it won't work for them since
they don't have that path. You can also use the `here` R package to
create paths. See examples of that below. I generally recommend the
`here` package.

**Rocky Mountain Spotted Fever (RMSF) Severity:**

A box plot of TITER1 values (log-transformed for better visualization)
provides insights into the variability of test results. RMSF severity
patterns are considered analogous to Alpha-Gal Syndrome, a topic of
interest, though not explicitly included in this dataset as it is not a
reportable disease in Georgia.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "rmsf_titer_boxplot_log.png")

# Display the image
knitr::include_graphics(img_path)

```

**Most Common Tick-Borne Diseases:**

The dataset is filtered to identify the five most frequently reported
tick-borne diseases. A bar chart visualizes their case counts, offering
an overview of significant public health concerns.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "disease_histogram.png")

# Display the image
knitr::include_graphics(img_path)
```

**Lyme Disease Analysis:**

Given its similarity to Alpha-Gal Syndrome (which is of interest to me
and the topic of my capstone), Lyme Disease cases are analyzed
separately. A count histogram compares positive, negative, and other
test outcomes, providing insights into testing patterns and diagnosis
rates.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "lyme_result_count_histogram2.png")

# Display the image
knitr::include_graphics(img_path)

```

**Geographic Distribution:**

A grouped bar chart maps case counts across Georgia counties,
categorized by disease type. Due to the density of information, this
visualization is refined to focus on the top five tick-borne diseases
and their distribution across the 12 counties with the highest confirmed
case counts.

```{r}
library(here)
library(knitr)

# Generate correct file paths
img_paths <- c(
  here::here("results", "figures", "disease_county_plot.png"),
  here::here("results", "figures", "top5_diseases_stacked_plot.png"),
  here::here("results", "figures", "top5_disease_county_plot.png")
)

# Display the images
knitr::include_graphics(img_paths)
```

To analyze disease burden trends and identify hotspot counties, a
stacked bar graph is generated to visualize case distribution across
different regions.

**Age Distribution Analysis**

Age distribution is assessed as part of the initial data cleaning
process, confirming that no cases in the cleaned dataset have missing
age values. Additionally, an exploratory analysis examines disease
counts by age group to identify potential patterns in disease
prevalence.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "disease_count_by_age.png")

# Display the image
knitr::include_graphics(img_path)
```

**Data Completeness Over Time**

To evaluate how data completeness has evolved, the dataset is processed
to extract onset date information. A visualization distinguishes between
blank and non-blank entries in the STATUS column over time, helping to
identify periods with high missingness. Spikes in blank values may
indicate data entry errors, system issues, or reporting gaps that should
be addressed before conducting further analysis.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "status_completeness_plot.png")

# Display the image
knitr::include_graphics(img_path)
```

**Timeline Analysis of Key Case Events** Given the frequent
loss-to-follow-up observed in tick-borne illness reporting, an analysis
is conducted to examine time gaps between critical dates, including:

-   **Onset date (DOO) to first lab test date (LABDATE1)**

-   **First lab test (LABDATE1) to second lab test (LABDATE2)**

-   **Second lab test (LABDATE2) to case update date (UPDATEDATE)**

These durations are summed to determine the total case timeline for each
record. The processed dataset is saved as a CSV file for further
analysis, and a formatted HTML table is generated to present the
extracted time intervals in a structured manner. Data Quality
Adjustments Upon reviewing the time calculations, inconsistencies were
identified where negative time values appeared between initial lab
results and follow-up tests. These errors, likely due to data entry
mistakes, made it impossible to determine the correct dates. To address
this, a filtering step was added to remove erroneous records where
follow-up test dates occurred before initial lab dates.

## Statistical analysis

{{< pagebreak >}}

## Basic statistical analysis

For the first key question related to the disease trends over time (in
years). In looking at the plot, it appears that there have been
increases over time were observed in most of the diseases. However, the
rate of increase was relatively slow as indicated by the lack of
statistical significance of these trends (p\>0.5) which may indicate an
attribution to the increase in reporting rather than an increase in
overall disease burden. In the full analysis and second iteration of
this, these trends will be compared year to year, comparing each year to
the 2018 standard.

+------------+------------+------------+------------+------------+
| term       | estimate   | std.error  | statistic  | p.value    |
|            |            |            |            |            |
| \<chr\>    | \<dbl\>    | \<dbl\>    | \<dbl\>    | \<dbl\>    |
+:===========+===========:+===========:+===========:+===========:+
| (          | -2         | 1          | -          | 0.1729390  |
| Intercept) | 3          | 573.052665 | 1.48025251 |            |
|            | 28.5151515 |            |            |            |
+------------+------------+------------+------------+------------+
| YEAR       | 1.1515152  | 0.777581   | 1.48089418 | 0.1727716  |
+------------+------------+------------+------------+------------+
| DISEAS     | 0.4242424  | 5.484565   | 0.07735207 | 0.9400358  |
| E          |            |            |            |            |
| BABESIOSIS |            |            |            |            |
+------------+------------+------------+------------+------------+
| DI         | 4.8484848  | 6.364775   | 0.76176850 | 0.4656907  |
| S          |            |            |            |            |
| EASEDENGUE |            |            |            |            |
+------------+------------+------------+------------+------------+
| DISEA      | 5.0000000  | 5.157889   | 0.96938889 | 0.3576794  |
| S          |            |            |            |            |
| EEHRLICHIA |            |            |            |            |
| C          |            |            |            |            |
| HAFFEENSIS |            |            |            |            |
+------------+------------+------------+------------+------------+
| DISEA      | -1.1515152 | 6.364775   | -          | 0.8604390  |
| S          |            |            | 0.18092002 |            |
| EEHRLICHIA |            |            |            |            |
| EWINGII    |            |            |            |            |
+------------+------------+------------+------------+------------+
| D          | 8.5606061  | 4.963745   | 1.72462659 | 0.1186814  |
| ISEASELYME |            |            |            |            |
| DISEASE    |            |            |            |            |
+------------+------------+------------+------------+------------+
| D          | 0.5757576  | 5.484565   | 0.10497781 | 0.9186958  |
| I          |            |            |            |            |
| SEASEROCKY |            |            |            |            |
| MOUNTAIN   |            |            |            |            |
| SPOTTED    |            |            |            |            |
| FEVER      |            |            |            |            |
+------------+------------+------------+------------+------------+
| D          | 8.2272727  | 5.593721   | 1.47080497 | 0.1754208  |
| ISEASEWEST |            |            |            |            |
| NILE (WNV) |            |            |            |            |
| INFECTION  |            |            |            |            |
+------------+------------+------------+------------+------------+

9 rows

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "trend_over_time_by_disease.png")

# Display the image
knitr::include_graphics(img_path)
```

The most interesting finding from the initial modeling was in relation
to the data completeness research question. In modeling the differences
across time in whether a final STATUS was inputted rather than left
blank, a general decrease was seen across the full 2018-2024 time scale.
Specifically, completion rates of the reporting form were approximately
50% in 2018 but dropped rapidly to 20% in 2019-- around which the rate
of completeness has hovered since 2019. However, in linear modeling, the
changes across time in the aggregate, p \> 0.05.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "completeness_over_time.png")

# Display the image
knitr::include_graphics(img_path)

```

+------------+------------+------------+------------+------------+
| term       | estimate   | std.error  | statistic  | p.value    |
|            |            |            |            |            |
| \<chr\>    | \<dbl\>    | \<dbl\>    | \<dbl\>    | \<dbl\>    |
+:===========+===========:+===========:+===========:+===========:+
| (          | 9          | 5          | 1.733920   | 0.1434648  |
| Intercept) | 935.267857 | 729.946482 |            |            |
+------------+------------+------------+------------+------------+
| YEAR       | -4.910714  | 2.835202   | -1.732051  | 0.1438108  |
+------------+------------+------------+------------+------------+

2 rows

On a positive note for the utilization of the reporting system and
trends towards better tick-bourne disease response, there has been
statistically significant decreases in the time between oneset and an
initial lab visit/interaction. In 2018, the average time between these
two occurrences ranged from approximately 150-520 days. With marked
decreases, the 2024 data indicates that patients with tick-bourne
disease occurrences may be interacting with providers or lab testing in
the same month, same week or even same day.

```{r}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "figures", "days_to_lab_over_time.png")

# Display the image
knitr::include_graphics(img_path)
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to
analyze your data and to produce meaningful figures, tables, etc. This
might again be code that is best placed in one or several separate R
scripts that need to be well documented. You want the code to produce
figures and data ready for display as tables, and save those. Then you
load them here.*

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

{{< pagebreak >}}

# References