### 2025-04-21 code/analysis-code/statistical-analysis-poverty-rate.qmd

```{r}
library(tidyverse)
library(tidymodels)
library(here)

# STEP 1: Summarize delay rate per county
delay_summary <- read_csv(here("data", "processed-data", "time_between_processeddata.csv"), show_col_types = FALSE) %>%
  mutate(COUNTY = str_to_title(COUNTY)) %>%
  group_by(COUNTY) %>%
  summarise(
    total_cases = n(),
    delay_cases = sum(TOTAL_DAYS > 30, na.rm = TRUE),
    delay_rate = delay_cases / total_cases
  ) %>%
  mutate(delay_rate_over_30 = as.factor(if_else(delay_rate > 0.3, 1, 0)))  # Binary outcome

# STEP 2: Load and prepare poverty rate data
poverty_data <- read_csv(here("data", "raw-data", "poverty_rate_by_county.csv"), show_col_types = FALSE) %>%
  transmute(
    COUNTY = str_remove(str_to_title(County), " County$"),
    POVERTY_RATE = as.numeric(POVERTY_RATE)
  )

# STEP 3: Merge datasets and clean
model_data <- delay_summary %>%
  left_join(poverty_data, by = "COUNTY") %>%
  drop_na(delay_rate_over_30, POVERTY_RATE)

# STEP 4: Define preprocessing recipe
model_recipe <- recipe(delay_rate_over_30 ~ POVERTY_RATE, data = model_data) %>%
  step_normalize(all_numeric_predictors())

# STEP 5: Define models
logistic_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

rf_model <- rand_forest(trees = 500) %>%
  set_engine("ranger") %>%
  set_mode("classification")

# STEP 6: Create workflows
logistic_workflow <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(logistic_model)

rf_workflow <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(rf_model)

# STEP 7: Cross-validation setup
set.seed(123)
cv_folds <- vfold_cv(model_data, v = 5, repeats = 3, strata = delay_rate_over_30)

# STEP 8: Fit models
logistic_res <- fit_resamples(logistic_workflow, resamples = cv_folds, metrics = metric_set(accuracy, roc_auc))
rf_res <- fit_resamples(rf_workflow, resamples = cv_folds, metrics = metric_set(accuracy, roc_auc))

# STEP 9: Collect performance metrics
logistic_metrics <- collect_metrics(logistic_res)
rf_metrics <- collect_metrics(rf_res)

# Display metrics
logistic_metrics
rf_metrics
```