### 2025-04-21 code/analysis-code/statistical-analysis.qmd

---
title: "Data Analysis for Raw Tick-Bourne Cases Data - 2019 to 2024"
author: "Hope Grismer"
date: "2025-03-10"
output: html_document
---

######################################
# analysis script
#
# This script loads the processed, cleaned data, does a simple analysis
# and saves the results to the results folder.

First, load all the libraries needed for tidymodels, multivariable GLM, and data visualization.

```{r}
library(tidymodels)  # For tidymodels framework
library(ggplot2)     # For data visualization
library(dplyr)       # For data manipulation
library(lubridate)   # For date manipulation
library(readr)       # For reading and writing CSV files
library(fs)          # For file system operations
library(broom)       # For tidying up GLM results
library(here)        # for data loading/saving
```

Then, load the cleaned datasets.

```{r}
# Load the cleaned dataset
cleaned_data2 <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Cleaned_Step2.rds"))
processed_time_data2 <- read_csv(here::here("data", "processed-data", "time_between_processeddata.csv"))

```

While the processing code creation involved some exploratory data analysis (EDA) already to help me determine what cleaning needed to be done, additional EDA helped find patterns in data to analyze further. See below.

```{r}

# Visualize time differences (e.g., DOO to LAB1)
ggplot(processed_time_data2, aes(x = DOO_to_LAB1)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Time from DOO to LAB1 (in days)", 
       x = "Days from DOO to LAB1", y = "Frequency")

```

######################################
# Data fitting/statistical analysis
#####################################

```{r}
# Path to the cleaned data (after saving it)
cleaned_data3 <- readRDS(here::here("data", "processed-data", "Tick-Bourne_Cases_Final_Cleaned.rds"))
```

```{r}
# 1. Disease trends over time:
# Extract YEAR from DOO (assuming DOO is a Date column)
cleaned_data3 <- cleaned_data3 %>%
  mutate(YEAR = as.integer(format(as.Date(DOO, format = "%Y-%m-%d"), "%Y")))

# Count cases grouped by DISEASE type and YEAR
disease_counts <- cleaned_data3 %>%
  group_by(DISEASE, YEAR) %>%
  summarise(CASE_COUNT = n(), .groups = "drop")

# Fit a linear model for yearly trend in cases (by DISEASE)
lmfit1 <- lm(CASE_COUNT ~ YEAR + DISEASE, data = disease_counts)
lmtable1 <- broom::tidy(lmfit1)
print(lmtable1)

# Save fit results
table_file1 <- here::here("results", "tables", "trend_over_time_by_disease.rds")
saveRDS(lmtable1, file = table_file1)

# Plot yearly trend by disease
plot1 <- ggplot(disease_counts, aes(x = YEAR, y = CASE_COUNT, color = DISEASE)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm") +
  labs(
    title = "Vector-Borne Diseases Over Time by Disease",
    x = "Year",
    y = "Number of Cases"
  ) +
  theme_minimal()

plot1

# Save plot
plot_file1 <- here::here("results", "supplemental-figures", "trend_over_time_by_disease.png")
ggsave(plot_file1, plot = plot1, width = 8, height = 6, dpi = 300)
```

```{r}
##Distribution Across Counties:
# Bar plot of cases by county
plot2 <- ggplot(cleaned_data3, aes(x = COUNTY, fill = DISEASE)) +
  geom_bar() +
  labs(
    title = "Cases by County",
    x = "County",
    y = "Number of Cases"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot2

# Save plot
plot_file2 <- here::here("results", "supplemental-figures", "cases_by_eachcounty.png")
ggsave(plot_file2, plot = plot2, width = 8, height = 6, dpi = 300)
```

```{r}
# 3. Completeness 
library(ggplot2)
library(dplyr)
library(lubridate)
library(broom)  # for tidying model results
library(readr)  # for reading CSV files

# Load the cleaned data (adjust path as needed)
statuscompleteness_summary_location <- here::here("data", "processed-data", "cleaned_statuscompleteness_summary.csv")
statuscompleteness_summary <- read_csv(statuscompleteness_summary_location)

# 1. Data Completeness Analysis Over Time:
# Summarize completeness over time
completeness_summary <- statuscompleteness_summary %>%
  mutate(YEAR = year(Month)) %>%
  group_by(YEAR) %>%
  summarise(COMPLETENESS = mean(Count) / sum(Count) * 100, .groups = "drop")  # Calculate completeness as percentage

# Linear Model: Analyze the relationship between YEAR and COMPLETENESS
lm_completeness <- lm(COMPLETENESS ~ YEAR, data = completeness_summary)

# Summarize the linear model results
lm_completeness_summary <- broom::tidy(lm_completeness)
print(lm_completeness_summary)

# 2. Create the completeness plot with linear trendline
plot3 <- ggplot(completeness_summary, aes(x = YEAR, y = COMPLETENESS)) +
  geom_line(color = "darkgreen") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +  # Add linear model trendline
  labs(
    title = "Data Completeness Over Time",
    x = "Year",
    y = "Completeness (%)"
  ) +
  theme_minimal()

# Print the plot
plot3

# Save the plot 
plot_file3 <- here::here("results", "supplemental-figures", "completeness_over_time.png")
ggsave(plot_file3, plot = plot3, width = 8, height = 6, dpi = 300)

```

```{r}
# 4. Timeline Between Key Dates:
# Fit a model for days between onset and lab over time
filtered_time_data <- cleaned_data3 %>%
  filter(!is.na(LABDATE1) & !is.na(DOO) & (LABDATE1 > DOO))

# Calculate the days between onset and lab result (assuming DATEONSET and LABDATE1 are in Date format)
filtered_time_data <- filtered_time_data %>%
  mutate(DAYS_TO_LAB = as.integer(LABDATE1 - DOO))

lmfit2 <- lm(DAYS_TO_LAB ~ YEAR, data = filtered_time_data)
lmtable2 <- broom::tidy(lmfit2)
print(lmtable2)

# Save fit results
table_file2 <- here::here("results", "tables", "days_to_lab_over_time.rds")
saveRDS(lmtable2, file = table_file2)

# Plot days between onset and lab over time
plot4 <- ggplot(filtered_time_data, aes(x = YEAR, y = DAYS_TO_LAB)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Days Between Onset and Lab Processing Over Time",
    x = "Year",
    y = "Days Between Onset and Lab"
  ) +
  theme_minimal()

plot4

# Save plot
plot_file4 <- here::here("results", "supplemental-figures", "days_to_lab_over_time.png")
ggsave(plot_file4, plot = plot4, width = 8, height = 6, dpi = 300)

```

```{r}
# 5. Relative Incidence of Disease Pathologies (2019–2024):
# Filter data for years 2019–2024
incidence_data <- cleaned_data3 %>%
  filter(YEAR >= 2019 & YEAR <= 2024)

# Group and calculate incidence by disease
incidence_summary <- incidence_data %>%
  group_by(YEAR, DISEASE) %>%
  summarise(INCIDENCE = n()) %>%
  ungroup()

# Plot relative incidence over time
plot5 <- ggplot(incidence_summary, aes(x = YEAR, y = INCIDENCE, color = DISEASE)) +
  geom_line(size = 1) +
  labs(
    title = "Relative Incidence of Disease Pathologies (2019–2024)",
    x = "Year",
    y = "Incidence"
  ) +
  theme_minimal()

plot5

# Save plot
plot_file5 <- here::here("results", "supplemental-figures", "relative_incidence_2019_2024.png")
ggsave(plot_file5, plot = plot5, width = 8, height = 6, dpi = 300)
```

```{r}
## ---- modeling_setup --------
library(tidymodels)

# Load the processed dataset
model_data <- readr::read_csv(here::here("data", "processed-data", "time_between_processeddata.csv"))

# Add the target variable and convert appropriate columns
model_data <- model_data %>%
  mutate(
    delay_over_threshold = if_else(TOTAL_DAYS > 30, 1, 0),  # Create binary target
    delay_over_threshold = as.factor(delay_over_threshold),  # Convert to factor
    DISEASE = as.factor(DISEASE),
    COUNTY = as.factor(COUNTY),
    SEX = as.factor(GENDER),
    AGE = as.numeric(AGE)  # Ensure numeric
  ) %>%
  drop_na(delay_over_threshold, AGE, SEX, COUNTY, DISEASE)  # Clean up

# Save the updated dataset with the new variable
write_csv(model_data, here::here("data", "processed-data", "time_between_processeddata.csv"))

```

To explore whether certain demographic or clinical variables could predict delayed follow-up in confirmed tick-borne disease cases, we created a binary outcome variable called delay_over_threshold, coded as 1 if the total time between disease onset and case closure exceeded 30 days, and 0 otherwise. This variable was calculated based on the sum of time intervals between key date fields (DOO, LABDATE1, LABDATE2, and UPDATEDATE), which had been previously cleaned and processed. Predictors selected for this analysis included age, sex, county, and disease type, as these are commonly available surveillance variables that may relate to access to care or follow-up procedures. All categorical predictors were converted into factor variables to prepare them for modeling.

```{r}
## ---- modeling_setup --------
library(tidymodels)

# Load the processed dataset
model_data <- readr::read_csv(here::here("data", "processed-data", "time_between_processeddata.csv"))

# Convert appropriate variables to factors
model_data <- model_data %>%
  mutate(
    delay_over_threshold = as.factor(delay_over_threshold),  # target variable
    DISEASE = as.factor(DISEASE),
    COUNTY = as.factor(COUNTY),
    SEX = as.factor(GENDER),
    AGE = as.numeric(AGE)  # just in case it's not numeric yet
  ) %>%
  drop_na(delay_over_threshold, AGE, SEX, COUNTY, DISEASE)  # Remove any rows with missing predictors

```

The data was randomly partitioned into training and testing sets using an 80/20 split, stratified by the outcome variable to maintain a balanced class distribution. The training set was used for model training and cross-validation, while the test set was reserved for final evaluation of model performance on unseen data.

```{r}
set.seed(123)
data_split <- initial_split(model_data, prop = 0.8, strata = delay_over_threshold)
train_data <- training(data_split)
test_data <- testing(data_split)

```

A recipe() was used to define preprocessing steps applied uniformly across all models. Categorical variables (such as county, disease, and sex) were one-hot encoded using dummy variables. Zero-variance predictors were removed, and the single numeric predictor, age, was normalized to ensure comparability across different model types.
```{r}
delay_recipe <- recipe(delay_over_threshold ~ AGE + SEX + COUNTY + DISEASE, data = train_data) %>%
  step_dummy(all_nominal_predictors()) %>%  # convert categorical predictors to dummy variables
  step_zv(all_predictors()) %>%  # remove zero-variance predictors
  step_normalize(all_numeric_predictors())  # standardize age
```

```{r}
logistic_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

tree_model <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")

rf_model <- rand_forest(trees = 500) %>%
  set_engine("ranger") %>%
  set_mode("classification")

```

For each model, a workflow was constructed using the tidymodels framework to combine the preprocessing recipe and the corresponding model specification. This approach ensured consistent preprocessing and simplified model training and evaluation.
```{r}
logistic_workflow <- workflow() %>%
  add_model(logistic_model) %>%
  add_recipe(delay_recipe)

tree_workflow <- workflow() %>%
  add_model(tree_model) %>%
  add_recipe(delay_recipe)

rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(delay_recipe)
```
To evaluate model performance and avoid overfitting, we implemented 5-fold cross-validation repeated 3 times. This resampling strategy provides a robust estimate of out-of-sample performance. Each model was trained and evaluated on multiple folds using two performance metrics: accuracy and the area under the ROC curve (AUC), the latter of which is especially informative for imbalanced or binary classification problems.
```{r}
set.seed(123)
cv_folds <- vfold_cv(train_data, v = 5, repeats = 3, strata = delay_over_threshold)

# Fit all models
logistic_res <- fit_resamples(logistic_workflow, resamples = cv_folds, metrics = metric_set(roc_auc, accuracy))
tree_res     <- fit_resamples(tree_workflow, resamples = cv_folds, metrics = metric_set(roc_auc, accuracy))
rf_res       <- fit_resamples(rf_workflow, resamples = cv_folds, metrics = metric_set(roc_auc, accuracy))

```
Cross-validation results for each model were collected and compared using mean accuracy and AUC. These metrics provided insight into the predictive quality of each approach. The model with the best combination of high accuracy and AUC was selected for final evaluation on the test set.
```{r}
collect_metrics(logistic_res)
collect_metrics(tree_res)
collect_metrics(rf_res)
```

```{r}
## ---- final_rf_evaluation --------
# Fit the final random forest model on training data and evaluate on test data
final_rf_fit <- rf_workflow %>%
  last_fit(split = data_split)

# Collect performance metrics
collect_metrics(final_rf_fit)

# Confusion matrix
collect_predictions(final_rf_fit) %>%
  conf_mat(truth = delay_over_threshold, estimate = .pred_class)

# ROC curve
collect_predictions(final_rf_fit) %>%
  roc_curve(truth = delay_over_threshold, .pred_1) %>%
  autoplot()

# Save ROC curve plot
roc_plot <- collect_predictions(final_rf_fit) %>%
  roc_curve(truth = delay_over_threshold, .pred_1) %>%
  autoplot() +
  theme_minimal() +
  labs(title = "ROC Curve for Random Forest Model")

ggsave(
  filename = here::here("results", "supplemental-figures", "roc_curve.png"),
  plot = roc_plot,
  width = 6, height = 5, dpi = 300
)
```