### 2025-04-21 products/manuscript/Manuscript.qmd

---
title: "Analysis of Tick-Bourne Disease Cases Reported to DPH Between 2019-2024"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

Authors

-   Hope Grismer $^{1,2,*, \land}$ (ORCID: 0000-0000-1234-5678)

Author affiliations

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Southeastern Center of Excellence in Vector-Bourne Diseases
    (SECVBD), Gainesville, FL 32611.

$\land$ Corresponding author: hope.grismer\@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Abstract

This project explores patterns in tick-borne disease cases reported to
the Georgia Department of Public Health (DPH) between 2019 and 2024. The
analysis involved extensive data cleaning to retain only confirmed and
positive cases, with additional processing to resolve inconsistencies in
timelines and reporting. Exploratory data analysis (EDA) was used to
examine disease distribution, demographic patterns, geographic hotspots,
and data completeness over time. Building on these insights, predictive
models were developed to assess whether patient and case-level
factors—such as age, disease type, and county—could help identify cases
at risk of delayed follow-up. This work theorectically supports improved
understanding of disease surveillance data and highlights areas for
strengthening tick-borne disease reporting and case management systems
in Georgia.

{{< pagebreak >}}

# Introduction

### 2.1 Background

In the past two decades, the study of tick-borne disease has gained
increasing attention due to rising incidence rates and expanding
geographic ranges of disease-carrying ticks. While early medical and
veterinary research in the 20th century identified key tick-borne
illnesses such as Rocky Mountain spotted fever and Lyme disease, recent
decades have seen an acceleration in the detection of new pathogens,
greater public health concern, and expanded surveillance efforts across
the globe. Researchers have documented the ecology and behavior of
ticks, explored the role of climate and land use in tick habitat
expansion, and analyzed patterns of disease incidence in both human and
animal populations. Despite the growing body of research, many knowledge
gaps remain, particularly in understanding co-infections, long-term
health effects, and effective prevention strategies. Tick-borne diseases
(TBDs) are primarily transmitted through the bite of infected
hard-bodied ticks (family Ixodidae), and include bacterial, viral, and
protozoan pathogens. Today, TBDs are recognized as a major public health
concern in both rural and suburban settings, with transmission risk
influenced by ecological, climatic, and sociocultural factors.

### 2.1.1 General Background Information

Tick-borne illnesses are diseases that humans can contract from the bite
of an infected tick. Ticks are small, blood-sucking arachnids that are
often found in wooded, grassy, or suburban areas. When a tick feeds on
an infected animal, it can acquire pathogens—such as bacteria, viruses,
or parasites—that can then be transmitted to humans. Examples of
tick-borne diseases include, but are not limited to, Lyme disease,
Anaplasmosis, and Babesiosis. These diseases are of increasing public
health concern, especially as the geographic range and seasonality of
ticks expand due to environmental and ecological changes.

### 2.1.2 Transmission and Reservoirs

Ticks act as both vectors and reservoirs for a variety of pathogens,
transmitting disease to humans and animals during blood meals. Most
tick-borne diseases are zoonotic, with small mammals, deer, and birds
acting as key reservoirs in natural cycles. Transmission to humans
typically occurs when a nymph or adult tick feeds for an extended
period, allowing the pathogen to transfer from the tick’s salivary
glands into the host’s bloodstream. Common tick-borne pathogens include
Borrelia burgdorferi (the agent of Lyme disease), Rickettsia rickettsii
(Rocky Mountain spotted fever), Anaplasma phagocytophilum, Ehrlichia
chaffeensis, and viruses such as Powassan and Heartland virus. Ticks
themselves are influenced by multiple environmental factors, including
temperature, humidity, vegetation, and host availability. Human exposure
risk is increased in areas with high host density, edge habitats between
forest and residential zones, and during seasons with peak tick
activity. Additionally, land use changes—such as suburban development,
deforestation, and agricultural expansion—can modify tick habitats and
affect pathogen dynamics.

### 2.1.3 Ecological and Health Impacts

The ecological consequences of tick-borne disease extend beyond human
health. Tick populations and their associated pathogens can alter
wildlife health and community composition. Infected animals may suffer
sublethal effects that reduce fitness, while high burdens of tick
parasitism can influence population dynamics, particularly in rodents
and ungulates. From a public health perspective, TBDs can cause a range
of acute and chronic symptoms in humans, from fever and fatigue to
long-term neurological and musculoskeletal complications.
Co-infections—when an individual is infected with more than one
pathogen—are increasingly recognized and may complicate diagnosis and
treatment. The public health burden is also shaped by disparities in
access to health care, awareness of TBDs, and availability of diagnostic
testing, which may result in underreporting or misdiagnosis. The costs
associated with tick-borne disease—including medical care, productivity
loss, and prevention efforts—represent a growing concern for health
systems, particularly in endemic areas.

### 2.1.4 Relevance

Tick-borne diseases have been documented in all regions of the United
States, with the southeastern, northeastern, and upper midwestern states
experiencing particularly high incidence rates. Climate models suggest
that warming temperatures and shifting precipitation patterns may expand
the range of ticks and lengthen their active season, potentially
increasing the geographic footprint and seasonality of tick-borne
disease transmission. In the southeastern U.S., for example, Amblyomma
americanum (the lone star tick) has expanded in abundance and geographic
range over the past decades, contributing to increased cases of
ehrlichiosis and the emergence of Alpha-Gal Syndrome, a red meat allergy
triggered by tick bites. Understanding the ecology, distribution, and
drivers of tick-borne disease is therefore crucial to predicting future
trends, identifying at-risk populations, and developing effective
prevention and response strategies.

# 3. The Data

# 3.1 Data Acquisition & Description

The data used in this analysis was obtained through my internship with
the Georgia Department of Health, in collaboration with the Centers for
Disease Control and Prevention (CDC). Specifically, it is derived from
the Nationally Notifiable Diseases Surveillance System (NNDSS) and
focuses on tick-borne illnesses. This data set provides key insights
into the trends, distribution, and characteristics of reported cases,
supporting public health efforts in disease monitoring and
prevention.2.1.5 Description of Data and Data Source

This data set includes all human cases reported to the Georgia
Department of Public Health (GA-DPH) between 2019 and 2024. It was
graciously provided by DPH professionals in the epidemiology department
and includes all available data for each case reported across the state
over this five-year period. Proper de-identification practices—such as
removal of first name, last name, and date of birth (DOB)—were followed.
The dataset includes variables such as disease type, patient
demographics (gender, age, district, state, county, and ZIP code),
medical details (incident ID, report ID, lab test dates, results,
serotype, and titers), and disease outcomes (death status, fever,
migraines, and administrative status). Reported diagnoses include
Toxoplasmosis, Dengue, Rocky Mountain Spotted Fever, Anaplasma
phagocytophilum, Malaria, Lyme Disease, West Nile Virus, Chikungunya,
LaCrosse Virus, Zika, Ehrlichia chaffeensis, and various forms of
Typhus. Some cases, based on adherence to specific case definitions and
subsequent lab results, are categorized as suspect, probable, confirmed,
or unconfirmed. By examining case reports, public health professionals
can better understand the prevalence, demographics, and regional
variations of these diseases. This analysis aims to further inform the
public health understanding of tick-borne disease in Georgia.

## 3.1.2 Questions and Hypotheses to Be Addressed

This project seeks to explore temporal, spatial, and epidemiologic
patterns of vector-borne disease (VBD) in Georgia using reported case
data from 2019–2024. Specifically, it aims to address the following
questions:

-   How complete is the data for vector-borne diseases, and how has data
    completeness evolved over time?

-   What is the timeline between key dates (onset, lab test, and date of
    entry), excluding same-day onset and lab results, and how does this
    timeline vary over time?

-   Can we predict whether a confirmed tick-borne disease case will
    result in a longer-than-typical diagnostic delay based on
    demographic and geographic features (county of origin of the case
    and Gini Index)?

    We hypothesized that tick-borne diseases would exhibit seasonal
    trends corresponding with tick life cycles, that higher incidence
    would be observed in areas with more natural and edge habitat, and
    that case counts would correlate with both environmental conditions
    and patterns of human interaction with tick habitats.

{{< pagebreak >}}

# 4. Methods

## 4.1 Workflow Summarized

-   Cleaned and prepared surveillance data on tick-borne disease cases
    reported in Georgia from 2019–2024, including filtering for
    confirmed positive cases and addressing issues with missing values
    and inconsistent timelines.

<!-- -->

-   Conducted exploratory data analysis (EDA) to examine temporal
    trends, geographic distribution, age patterns, diagnostic test
    outcomes, and data completeness, using visualizations to inform
    hypotheses and modeling decisions.

-   Developed and evaluated multiple predictive models to assess whether
    demographic and case-specific factors could explain delays in case
    follow-up, using cross-validation and test/train splitting to assess
    model quality and performance.

## 4.2 Uncleaned Data

The data contains 6255 entries representing discrete cases of
tick-bourne diseases. For each case, some or all of the following data
variables were collected: disease type, patient demographics (gender,
age, district, state, county, and ZIP code), medical details (incident
ID, report ID, lab test dates, results, serotype, and titers), and
disease outcomes (death status, fever, migranes, and administrative
status). For some cases, specific variables were not reported and,
therefore, are reported as blank spaces in the data set.

## 4.3 Initial Data Cleaning

The data cleaning and preprocessing workflow begins with loading
essential R packages for data handling (dplyr, tidyr, readxl), summary
visualization (skimr), and file path management (here). The raw dataset,
stored in an Excel file, is imported into R for structured processing.
To assess data quality and structure, we utilize functions such as
glimpse(), head(), and summary(), which provide an overview of column
names, data types, and potential inconsistencies like missing or
unexpected values. Additionally, skimr::skim() offers a detailed
summary, highlighting numerical distributions and missing data patterns.

The initial cleaning step removes records where the STATUS field is
either missing or not classified as "CONFIRMED," ensuring that only
officially verified cases are included. The dataset is then refined
further by retaining only cases with a confirmed "positive" test result,
excluding records with missing, blank, or ambiguous values (e.g.,
"inconclusive"). This process initially removed 15 cases that did not
meet the case definition in Georgia, leaving 95 valid records. Upon
further review, filtering criteria were adjusted to ensure all relevant
positive cases were retained while systematically excluding non-positive
values such as "Negative," "Equivocal," "Not detected," and other
irrelevant categories. This refinement resulted in a more accurate
dataset and a clearer summary table of key case information. The cleaned
dataset is saved at each stage to maintain version control and
reproducibility.

#### 4.3.1 Looked for Missing Data & General Trends

-   Excluded cases with a missing or non-“CONFIRMED” STATUS to retain
    only confirmed cases.
-   Refined the dataset by retaining only confirmed “positive” test
    results, excluding missing or ambiguous results, leaving 95 valid
    cases.
-   Additional filtering ensured correct identification of positive
    cases while excluding non-positive values.

#### Addressed Data Quality Issues

-   Identified negative time values between lab test dates, indicating
    data entry errors.
-   Implemented a filtering step to remove records with invalid date
    sequences.

## 4.4 Exploratory/Descriptive Analysis

#### 4.4.1 Identified Disease Burden Trends & Hotspot Counties

-   Generated a histogram to visualize the distribution of tick-borne
    diseases in Georgia between 2019 and 2024.
-   Created a bar chart highlighting the five most common diseases.
-   Specifically analyzed Lyme Disease cases, focusing on test results.

#### 4.4.2 Checked Age Distribution

-   Verified that no cases had missing age values in the cleaned
    dataset.
-   Conducted an exploratory analysis of disease counts by age group.

#### 4.4.3 Evaluated Data Completeness Over Time

-   Extracted onset date information to track missing data trends.
-   Created a plot distinguishing between blank and non-blank STATUS
    entries over time to identify periods of high missingness,
    indicating possible data entry errors or reporting gaps.

#### 4.4.4 Analyzed Timeline Between Key Case Events

-   Calculated time differences for:
    -   Onset date (DOO) to first lab test date (LABDATE1)
    -   First lab test (LABDATE1) to second lab test (LABDATE2)
    -   Second lab test (LABDATE2) to case update date (UPDATEDATE)
-   Summed these values to determine the total case duration.
-   Saved the processed dataset as a CSV file for further analysis.
-   Created a formatted HTML table to present the extracted time
    intervals.

## 4.5 EDA Revelant Takeaways

#### 4.5.1 Most Common Tick-Borne Diseases

The dataset is filtered to identify the five most frequently reported
tick-borne diseases. A bar chart visualizes their case counts, offering
an overview of significant public health concerns.

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "supplemental-figures", "disease_histogram.png")

# Display the image
knitr::include_graphics(img_path)
```

#### 4.5.2 Geographic Distribution

A grouped bar chart maps case counts across Georgia counties,
categorized by disease type to analyze disease burden trends and
identify hotspot counties. Due to the density of information, this
visualization is refined to focus on the top five tick-borne diseases
and their distribution across the 12 counties with the highest confirmed
case counts.

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file paths
img_paths <- c(
  here::here("results", "supplemental-figures", "disease_county_plot.png"),
  here::here("results", "supplemental-figures", "top5_diseases_stacked_plot.png"),
  here::here("results", "supplemental-figures", "top5_disease_county_plot.png")
)

# Display the images
knitr::include_graphics(img_paths)
```

#### 4.5.3 Data Completeness Over Time

To evaluate how data completeness has evolved, the dataset is processed
to extract onset date information. A visualization distinguishes between
blank and non-blank entries in the STATUS column over time, helping to
identify periods with high missingness. Spikes in blank values may
indicate data entry errors, system issues, or reporting gaps that should
be addressed before conducting further analysis.

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "supplemental-figures", "status_completeness_plot.png")

# Display the image
knitr::include_graphics(img_path)
```

For more of my initial exploratory data analysis, please see the
[Supplemental Figures]{.underline} section at the bottom of the report.

#### 4.5.4 Timeline Analysis of Key Case Events

Given the frequent loss-to-follow-up observed in tick-borne illness
reporting, an analysis is conducted to examine time gaps between
critical dates, including:

-   Onset date (DOO) to first lab test date (LABDATE1)

-   First lab test (LABDATE1) to second lab test (LABDATE2)

-   Second lab test (LABDATE2) to case update date (UPDATEDATE)

These durations are summed to determine the total case timeline for each
record. The processed dataset is saved as a CSV file for further
analysis, and a formatted HTML table is generated to present the
extracted time intervals in a structured manner. Data Quality
Adjustments Upon reviewing the time calculations, inconsistencies were
identified where negative time values appeared between initial lab
results and follow-up tests. These errors, likely due to data entry
mistakes, made it impossible to determine the correct dates. To address
this, a filtering step was added to remove erroneous records where
follow-up test dates occurred before initial lab dates.

## 4.6 Statistical Analysis

To evaluate whether diagnostic delays could be predicted from case-level
characteristics, a binary outcome variable (`delay_over_threshold`) was
created to indicate whether the total time from symptom onset to final
update exceeded 30 days. Predictor variables included patient age, sex,
county of residence, and disease type. Age was normalized, and
categorical variables were one-hot encoded.

Three classification models—logistic regression, decision tree, and
random forest—were trained using the `tidymodels`framework. Model
performance was assessed using 5-fold cross-validation repeated three
times. Metrics included accuracy and area under the receiver operating
characteristic curve (AUC).

Both basic bivariate analysis and statistical modeling was conducted.
For the statistical modeling, three classification models were trained
and evaluated using 5-fold cross-validation repeated 3 times.
Performance was assessed using accuracy and area under the ROC curve
(AUC).

# 5. Results

## 5.1 Basic statistical analysis

For the first key question related to the disease trends over time (in
years). In looking at the plot, it appears that there have been
increases over time were observed in most of the diseases. However, the
rate of increase was relatively slow as indicated by the lack of
statistical significance of these trends (p\>0.5) which may indicate an
attribution to the increase in reporting rather than an increase in
overall disease burden. In the full analysis and second iteration of
this, these trends will be compared year to year, comparing each year to
the 2018 standard.

+-----------------+-----------+-----------+-----------+-----------+
| term            | estimate  | s         | statistic | p.value   |
|                 |           | td.error  |           |           |
| \<chr\>         | \<dbl\>   |           | \<dbl\>   | \<dbl\>   |
|                 |           | \<dbl\>   |           |           |
+:================+==========:+==========:+==========:+==========:+
| ( Intercept)    | -2 3 2    | 1 57      | 1         | 0.1729390 |
|                 | 8.5151515 | 3.052665  | .48025251 |           |
+-----------------+-----------+-----------+-----------+-----------+
| YEAR            | 1.1515152 | 0.777581  | 1         | 0.1727716 |
|                 |           |           | .48089418 |           |
+-----------------+-----------+-----------+-----------+-----------+
| BABESIOSIS      | 0.4242424 | 5.484565  | 0         | 0.9400358 |
|                 |           |           | .07735207 |           |
+-----------------+-----------+-----------+-----------+-----------+
| DENGUE          | 4.8484848 | 6.364775  | 0         | 0.4656907 |
|                 |           |           | .76176850 |           |
+-----------------+-----------+-----------+-----------+-----------+
| EHRLICHIA       | 5.0000000 | 5.157889  | 0         | 0.3576794 |
| CHAFFEENSIS     |           |           | .96938889 |           |
+-----------------+-----------+-----------+-----------+-----------+
| EHRLICHIA       | -         | 6.364775  | 0         | 0.8604390 |
| EWINGII         | 1.1515152 |           | .18092002 |           |
+-----------------+-----------+-----------+-----------+-----------+
| LYME DISEASE    | 8.5606061 | 4.963745  | 1         | 0.1186814 |
|                 |           |           | .72462659 |           |
+-----------------+-----------+-----------+-----------+-----------+
| ROCKY MOUNTAIN  | 0.5757576 | 5.484565  | 0         | 0.9186958 |
| SPOTTED FEVER   |           |           | .10497781 |           |
+-----------------+-----------+-----------+-----------+-----------+
| WEST NILE (WNV) | 8.2272727 | 5.593721  | 1         | 0.1754208 |
| INFECTION       |           |           | .47080497 |           |
+-----------------+-----------+-----------+-----------+-----------+

9 rows

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "supplemental-figures", "trend_over_time_by_disease.png")

# Display the image
knitr::include_graphics(img_path)
```

The most interesting finding from the initial modeling was in relation
to the data completeness research question. In modeling the differences
across time in whether a final STATUS was inputted rather than left
blank, a general decrease was seen across the full 2018-2024 time scale.
Specifically, completion rates of the reporting form were approximately
50% in 2018 but dropped rapidly to 20% in 2019-- around which the rate
of completeness has hovered since 2019. However, in linear modeling, the
changes across time in the aggregate, p \> 0.05, meaning observed
differences from year to year were not statistically significant.

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "supplemental-figures", "completeness_over_time.png")

# Display the image
knitr::include_graphics(img_path)

```

+----------+----------+----------+----------+----------+
| term     | estimate | s        | s        | p.value  |
|          |          | td.error | tatistic |          |
| \<chr\>  | \<dbl\>  |          |          | \<dbl\>  |
|          |          | \<dbl\>  | \<dbl\>  |          |
+:=========+=========:+=========:+=========:+=========:+
| ( I      | 9 9      | 5 7      | 1.733920 | 0        |
| n        | 3        | 2        |          | .1434648 |
| tercept) | 5.267857 | 9.946482 |          |          |
+----------+----------+----------+----------+----------+
| YEAR     | -        | 2.835202 | -        | 0        |
|          | 4.910714 |          | 1.732051 | .1438108 |
+----------+----------+----------+----------+----------+

2 rows

On a positive note for the utilization of the reporting system and
trends towards better tick-borne disease response, there has been
statistically significant decreases in the time between onset and an
initial lab visit/interaction. In 2018, the average time between these
two occurrences ranged from approximately 150-520 days. With marked
decreases, the 2024 data indicates that patients with tick-borne disease
occurrences may be interacting with providers or lab testing in the same
month, same week or even same day.

```{r, echo=FALSE}
library(here)
library(knitr)

# Generate correct file path
img_path <- here::here("results", "supplemental-figures", "days_to_lab_over_time.png")

# Display the image
knitr::include_graphics(img_path)
```

## 5.2 Further Statistical Analysis

To assess whether delays in disease follow-up could be predicted based
on demographic or case-specific factors, a binary classification
variable (`delay_over_threshold`) was created. This variable indicates
whether the total case duration—from symptom onset to final
update—exceeded 30 days. The outcome was treated as a binary factor (1 =
delay \> 30 days, 0 = delay ≤ 30 days) and used as the target variable
for modeling.

Predictors included:

-   AGE: Patient's age at time of case entry

-   SEX: Reported gender

-   COUNTY: County of residence

-   DISEASE: Tick-borne disease type (e.g., Ehrlichiosis, Lyme disease,
    RMSF)

These predictors were selected based on their relevance to health access
and geographic exposure patterns. All categorical variables were
converted into dummy variables for modeling, and the age variable was
normalized. Three classification models were trained and evaluated using
the `tidymodel`framework:

1.  Logistic Regression

2.  Decision Tree

3.  Random Forest

To evaluate model performance, 5-fold cross-validation was repeated 3
times on the training dataset. Model quality was assessed using both
accuracy (overall correct classification rate) and area under the ROC
curve (AUC), which measures the model's ability to discriminate between
delayed and non-delayed cases.

| Model               | Accuracy | ROC AUC |
|---------------------|----------|---------|
| Logistic Regression | 72.8%    | 0.65    |
| Decision Tree       | 77.8%    | 0.61    |
| Random Forest       | 78.1%    | 0.68    |

The logistic regression model provided an interpretable baseline with
modest discriminatory power (AUC = 0.65). The decision tree achieved
higher accuracy but lower AUC, indicating possible overfitting or
limited generalizability. The random forest model outperformed the other
two in both metrics, achieving 78.1% accuracy and an AUC of 0.68, and
was selected as the best-performing model. Given these results, the
random forest model was selected for final evaluation on the test data
set.

```{r, echo=FALSE}
library(here)
library(knitr)

knitr::include_graphics(here::here("results", "supplemental-figures", "roc_curve.png"))

```

## 5.3 Adding in Poverty Rate Data by County

Now, I was curious. Reflecting back on the EDA analysis and the stark
differences in county and cases, if the modeling of delay would better
predicted if we merged county-level poverty rate data. This poverty rate
data is reasonable, potential structural determinant of diagnostic delay
due to persistent healthcare inequities. This variable was included in
model training alongside demographic and disease characteristics.
Preprocessing steps were updated accordingly, and models were
re-evaluated using the same cross-validation and test evaluation
procedures to assess any added predictive value.

County-level poverty rate was evaluated as a predictor of whether a
county had a high proportion of cases with diagnostic delays exceeding
30 days. A random forest classification model using only poverty rate as
a predictor achieved 88.4% accuracy, but its ability to discriminate
between delayed and non-delayed counties was poor (ROC AUC = 0.46). This
suggests that although the model often guessed the majority class
correctly, it lacked the ability to meaningfully differentiate high-risk
counties.

Similarly, a linear regression model found no statistically significant
association between poverty rate and delay rate (p = 0.87, β = 0.0049).
The random forest regression model explained less than 4% of the
variance in delay rates (R² = 0.038) with a relatively high prediction
error (RMSE = 0.37).

+----------------+----------------+----------------+----------------+
|                |                |                |                |
+----------------+----------------+----------------+----------------+
|                |                |                |                |
+----------------+----------------+----------------+----------------+

{{< pagebreak >}}

# 6. Discussion

## 6.1 Summary and Interpretation

The random forest model for the modeling of delay as an outcome all
predictors selected from cross-validation was refitted to the full
training data and evaluated on the reserved 20% test dataset to assess
generalizability. On this unseen data, the model achieved an accuracy of
74.7%, meaning nearly three-quarters of tick-borne disease cases were
correctly classified as having experienced delayed (\>30 day) or
non-delayed follow-up. The model’s ability to discriminate between
delayed and non-delayed cases was moderate, with an area under the
receiver operating characteristic (ROC) curve of 0.688, closely
mirroring performance observed during cross-validation.

The ROC curve showed consistent separation between sensitivity and
1-specificity, further supporting model validity. These results suggest
that case characteristics such as age, sex, county, and disease type
contain meaningful predictive information regarding follow-up delays in
surveillance data.Such insights can inform improvements in case
monitoring, triaging, and outreach in counties or patient groups with
higher risk for delays.

County-level poverty rate was evaluated as a predictor of the proportion
of cases with diagnostic delays exceeding 30 days. A linear regression
model indicated no statistically significant relationship between
poverty rate and delay rate (p = 0.87). The estimated effect size was
minimal (β = 0.0049), and the 95% confidence interval included zero.

A random forest regression model using the same predictor resulted in
low predictive performance. The R² was 0.038, indicating that poverty
rate explained less than 4% of the variance in delay rate across
counties. The RMSE was 0.37, reflecting high residual error in
predicting delay proportions.

In comparison, the earlier classification models using individual-level
predictors (e.g., age, sex, disease type, county) performed considerably
better in both accuracy and AUC, suggesting that county-level poverty
alone does not substantially contribute to explaining or predicting
diagnostic delays in this dataset.

### What this means:

-   The poverty-only model had higher accuracy, but a lower ROC AUC.

-   Accuracy tells us how often the model got it right — so poverty
    alone seems to classify “delay” correctly most of the time.

-   However, the low ROC AUC suggests that the model struggles to rank
    cases by risk — in other words, it doesn’t distinguish well between
    high and low risk of delay.

-   The all-predictors model performed more balanced, with a decent
    accuracy and a stronger ROC AUC, meaning it did a better
    job discriminating between delayed and non-delayed cases.

+----------------+----------------+----------------+----------------+
| Model Type     | Predictors     | Accuracy       | ROC AUC        |
|                | Used           |                |                |
+================+================+================+================+
| Random Forest  | Age, Sex,      | 78.1%          | 0.68           |
|                | County,        |                |                |
|                | Disease        |                |                |
+----------------+----------------+----------------+----------------+
| Random         | Poverty Rate   | 88.4%          | 0.46           |
| Forest -       | By County      |                |                |
| Poverty Rate   |                |                |                |
| Only           |                |                |                |
+----------------+----------------+----------------+----------------+

## 6.2 Strengths and Limitations

This analysis leveraged a comprehensive dataset spanning six years and
multiple counties across Georgia, allowing for exploration of both
individual- and community-level patterns in diagnostic delay for
tick-borne disease cases. The geographic breadth of the data is a key
strength, enabling comparisons across rural and urban areas and
supporting more generalizable insights. Additionally, the use of several
modeling strategies—including logistic regression, decision trees, and
random forests—provided a well-rounded evaluation of predictor
importance and model performance. The inclusion of both cross-validation
and a holdout test set added rigor and reduced the risk of overfitting.

Another limitation is the use of county-level poverty rate as a proxy
for individual socioeconomic status (SES). While poverty data at the
county level offers a broad socioeconomic context, it can mask
significant within-county income disparities. In counties with wide
economic variation, aggregate poverty rates may not accurately represent
the lived experience or access to healthcare of individuals affected by
tick-borne disease.

Furthermore, individual exposure risk is often tightly linked to SES.
For instance, people in outdoor labor roles—such as landscaping,
farming, or construction—may face greater tick exposure than those
working indoors, regardless of the county’s average income level. These
more granular, individual-level occupational or behavioral factors were
not captured in this dataset, which limits the precision of predictive
modeling based solely on geography and aggregated socioeconomic
indicators.

However, several limitations should be acknowledged. Inconsistencies in
reporting across cases and counties required extensive data cleaning,
which ultimately reduced the dataset to a much smaller number of usable
observations. This constrained sample size may have limited model
performance and generalizability. The dataset also lacked clinical and
healthcare access variables—such as symptom severity, insurance status,
or provider responsiveness—which likely influence diagnostic delays.
Without these contextual details, the models may have missed key drivers
of delay. Finally, the poverty rate was measured at the county level,
which may not reflect intra-county disparities or the lived experience
of patients.

# 7. Conclusions

-   Individual-level characteristics such as age, sex, disease type, and
    county were moderately effective in predicting whether a tick-borne
    disease case experienced a diagnostic delay of over 30 days. The
    best-performing model (random forest) achieved 78.1% accuracy and an
    AUC of 0.68, indicating meaningful—though not perfect—predictive
    value.

-   County-level poverty rate alone was not a strong predictor of
    diagnostic delay. Both linear regression and random forest
    regression showed weak associations and low explanatory power,
    suggesting that broader socioeconomic indicators may not capture the
    nuanced drivers of individual-level follow-up delays.

-   Data quality and completeness were major challenges: extensive
    cleaning was required, and missing or inconsistent entries
    significantly reduced the number of usable cases. This limitation
    underscores the need for more standardized reporting practices in
    public health surveillance systems.

{{< pagebreak >}}

# Supplemental Figures

A full set of exploratory and supplemental visualizations—used to guide
data cleaning, quality checks, and hypothesis generation—are available
in a separate folder.

Please refer to the following file for all supplemental figures:

File: `supplemental-figures.zip`

This archive includes: - Disease distribution histograms - County-level
maps - Test result breakdowns - Titer plots - Age distribution and data
completeness trends that I did in the exploratory data analysis but
opted to not include in my analysis per my reviewers' recommendations
that I narrow my scope!

# References